Computing the {\em least-commitment} action model from state observations

### 2018-12-19 doc/ijcai-19-conformant/conformant-learning.tex
first updates added after discussing the points made by blai

\subsection{Computing the {\em least-commitment} model via classical planning}
Inspired by the {\em classical planning compilation} $K_{s_0}$ for conformant planning~\cite{palacios-conformant-JAIR09}, this section shows that we can build a {\em classical planning problem} $P=\tup{F',A',I',G'}$ whose solution induces the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$: 
\begin{itemize}
\item The set of fluents $F'$ extends $F$ with two new sets of fluents:
      \begin{itemize}
      \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
      \item Fluents {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$}, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} encoding the {\em knowledge level} representation of the space of possible \strips\ action models.
      \end {itemize}
\item The set of actions $A'$ contains now actions of three different kinds:
\begin{itemize}
      \item Actions for {\em committing} {\tt\small pre\_e\_$\xi$} to a positive/negative value. Similar actions are also defined for {\em committing} {\tt\small eff\_e\_$\xi$} to a positive/negative value but the value of {\tt\small eff\_e\_$\xi$} can only be committed once the value of the corresponding {\tt\small pre\_e\_$\xi$} is committed (i.e. once either $\mathsf{K\pre\_e\_\xi}$ or $\mathsf{K\neg\pre\_e\_\xi}$ holds in the current state).
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{commit\top\_pre\_e\_\xi})=&\{mode_{commit}, \\
&\mathsf{\neg K\pre\_e\_\xi}, \mathsf{\neg K\neg\pre\_e\_\xi}\},\\
\cond(\mathsf{commit\top\_pre\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\pre\_e\_\xi}\}.\\\\
\hspace*{7pt}\pre(\mathsf{commit\bot\_pre\_e\_\xi})=&\{mode_{commit}, \\
&\mathsf{\neg K\pre\_e\_\xi},\mathsf{\neg K\neg\pre\_e\_\xi}\},\\
\cond(\mathsf{commit\bot\_pre\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\neg\pre\_e\_\xi}\}.
\end{align*}
\end{small}

@@

      \item {\em Editable} actions whose semantics is given by the value of the {\em knowledge level} fluents ({\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$}, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$}) at the current state. Figure~\ref{fig:editable} shows the PDDL encoding of an {\em editable} {\tt\small stack(?v1,?v2)} schema. This editable schema behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional} when the set of fluents {\tiny\tt (Kpre\_holding\_v1\_stack) (Kpre\_clear\_v2\_stack) (Keff\_holding\_v1\_stack) (Keff\_clear\_v2\_stack) (Keff\_clear\_v1\_stack) (Keff\_handempty\_stack) (Keff\_on\_v1\_v2\_stack)} hold at the current state. Formally, given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is:
\begin{small}  
\begin{align*}
\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{\neg K\neg\pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
\cond(\mathsf{editable_{\xi}})=&\{K\pre\_e\_\xi, \neg K\neg\eff\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
&\{K\neg\pre\_e\_\xi, \neg K\neg\eff\_e\_\xi\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
\end{align*}
\end{small}

>>>>

\subsection{Computing the {\em least-commitment} model via classical planning}
Inspired by the {\em classical planning compilation} $K_{s_0}$ for conformant planning~\cite{palacios-conformant-JAIR09}, this section shows that we can build a {\em classical planning problem} $P=\tup{F',A',I',G'}$ whose solution induces the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$: 
\begin{itemize}
\item The set of fluents $F'$ extends $F$ with two new sets of fluents:
      \begin{itemize}
      \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
      \item The {\em knowledge level} fluents {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$}, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} encoding the space of possible {\em partially specified} action models.
      \end {itemize}
\item The set of actions $A'$ contains now actions of three different kinds:
\begin{itemize}
      \item Actions for {\em committing} {\tt\small pre\_e\_$\xi$} to a positive/negative value. Similar actions are also defined for {\em committing} {\tt\small eff\_e\_$\xi$} to a positive/negative value but the value of {\tt\small eff\_e\_$\xi$} can only be committed once the value of the corresponding {\tt\small pre\_e\_$\xi$} is committed (i.e. once either $\mathsf{K\pre\_e\_\xi}$ or $\mathsf{K\neg\pre\_e\_\xi}$ holds in the current state).
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{commit\top\_pre\_e\_\xi})=&\{mode_{commit}, \\
&\mathsf{\neg K\pre\_e\_\xi}, \mathsf{\neg K\neg\pre\_e\_\xi}\},\\
\cond(\mathsf{commit\top\_pre\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\pre\_e\_\xi}\}.\\\\
\hspace*{7pt}\pre(\mathsf{commit\bot\_pre\_e\_\xi})=&\{mode_{commit}, \\
&\mathsf{\neg K\pre\_e\_\xi},\mathsf{\neg K\neg\pre\_e\_\xi}\},\\
\cond(\mathsf{commit\bot\_pre\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\neg\pre\_e\_\xi}\}.
\end{align*}
\end{small}

@@

      \item {\em Editable} actions whose semantics is given by the value of the {\em knowledge level} fluents ({\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$}, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$}) at the current state. Figure~\ref{fig:editable} shows the PDDL encoding of an {\em editable} {\tt\small stack(?v1,?v2)} schema. This editable schema behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional} when the set of fluents {\tiny\tt (Kpre\_holding\_v1\_stack) (Kpre\_clear\_v2\_stack) (Keff\_holding\_v1\_stack) (Keff\_clear\_v2\_stack) (Keff\_clear\_v1\_stack) (Keff\_handempty\_stack) (Keff\_on\_v1\_v2\_stack)} hold at the current state as well as the remaining {\tt\small $K\neg\pre\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} for all the preconditions and effects that are not part of the actual {\tt\small stack(?v1,?v2)} schema. Formally, given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is:
\begin{small}  
\begin{align*}
\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{\neg K\neg\pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
\cond(\mathsf{editable_{\xi}})=&\{K\pre\_e\_\xi, \neg K\neg\eff\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
&\{K\neg\pre\_e\_\xi, \neg K\neg\eff\_e\_\xi\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
\end{align*}
\end{small}
